Report on Lab 1 \& Lab 2 in EDA282, Parallel Computer Organization and Design

### 2014-10-11 lab/report/include/introduction.tex

\section{Introduction}
\label{sec:int}
These laboratory assignments aim to increase our knowledge of different cache coherence protocols. The goal is to simulate cache coherence protocols on a modeled cache coherent multi-core system using diverse benchmarks. Three cache coherence protocols are targeted; MSI, MESI, and MESI-MG, where the last protocol is an modified version of MESI optimized for migratory sharing. The system model have a shared Last Level Cache (LLC), and each processor have a private L1 cache. Snooping based write-invalidate cache coherence protocols are used by the private L1 caches to ensure coherence, and a simple bus stand for the communication between LLC and private caches. The system is modeled by the Pin tool MultiCacheSim and Pin.

>>>>

\section{Introduction}
\label{sec:int}
These laboratory assignments aim to increase our knowledge of different cache coherence protocols. The goal is to simulate cache coherence protocols (CCP) on a modeled cache coherent multi-core system using diverse benchmarks. Three cache coherence protocols are targeted; MSI, MESI, and MESI-MG, where the last protocol is an modified version of MESI optimized for migratory sharing. The system model have a shared Last Level Cache (LLC), and each processor have a private L1 cache. Snooping based write-invalidate cache coherence protocols are used by the private L1 caches to ensure coherence, and a simple bus stand for the communication between LLC and private caches. The system is modeled by the Pin tool MultiCacheSim and Pin. In listing~\ref{lst:setup} is the command used for simulating a CCP on the modeled system with a chosen benchmark.

\begin{lstlisting}[label=lst:setup, basicstyle=\scriptsize, caption={Command used for simulation.}]
#> ./pin -mt -t ./source/tools/MultiCacheSim/obj-intel64/mcs.so -csize 65536 -bsize 64 -assoc 8 -llc -llcsize 4194304 -llcbsize 64 -llcassoc 16 -numcaches 8 -protos ./source/tools/MultiCacheSim/obj-intel64/mcs_<CCP>.so -- <benchmark> [#threads,..]
\end{lstlisting}

### 2014-10-11 lab/report/include/lab1.tex
added intro

\section{Lab 1}
\label{sec:lab1}
\todo[inline]{Introduction, one paragraph}

\subsection{Task 1}
\todo[inline]{Explain the benchmarks}

\subsection{Task 2}
\todo[inline]{
Compare MSI and MESI.
When is having the E state beneficial (over baseline MSI)?
How does the behavior change when the dataset size changes?}

\subsection{Task 3}
\todo[inline]{
Compare MESI and MESI+migratory.
Why is one better than the other?
How does the behavior change when the dataset size changes?}

>>>>

\section{Lab 1}
\label{sec:lab1}
\todo[inline]{little drafty}
In this laboratory assignment, we evaluate overheads in cache coherence protocols using microbenchmarks. Three diverse benchmarks, described in \ref{sec:lab11}, are used to compare the coherence protocols. In \ref{sec:lab12}, MSI and MESI is compared and in \ref{sec:lab13}, MESI and MESI-MG is compared.

\todo[inline]{Add working size set}

\subsection{Task 1}
\label{sec:lab11}
Three different microbenchmarks are used with various prominent communication patterns. Here follows a description of each benchmark.
\subsubsection*{Benchmark 1 - Producer consumer}
In this benchmark, one thread increments the values in a memory array, whereas the other threads read these values and accumulate the read values in a private variable. However, it is not ensured that the incrementing thread increments the values before the other threads reads, and thus can this happen in any order. The memory accesses is done in mutual exclusion using a lock, and before a new lap of increments \& reads there is a barrier.

\subsubsection*{Benchmark 2 - No sharing}
The memory array is divided up equally among the threads and a thread only increments the values in it's assigned array elements, thus is the memory array not shared. A barrier is placed before a new incrementation lap of the assigned array elements begins.

\subsubsection*{Benchmark 3 - All shared}
All threads increments every element in the memory array.

\subsection{Task 2}
\label{sec:lab12}
\todo[inline]{
Compare MSI and MESI.
When is having the E state beneficial (over baseline MSI)?
How does the behavior change when the dataset size changes?}

\subsection{Task 3}
\label{sec:lab13}
\todo[inline]{
Compare MESI and MESI+migratory.
Why is one better than the other?
How does the behavior change when the dataset size changes?}
